\documentclass[sigconf,anonymous]{acmart}

% Remove ACM-specific metadata for draft
\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}

\usepackage{booktabs}
\usepackage{amsmath}
\let\Bbbk\relax  % Fix conflict between amssymb and newtx
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{subcaption}
\usepackage{multirow}

% Required ACM metadata
\keywords{cryptocurrency, cross-sectional prediction, distributional forecasting, Transformer, Student-t}

\begin{document}

% ═══════════════════════════════════════════════════════════════
% TITLE
% ═══════════════════════════════════════════════════════════════

\title{Phantom: Distributional Forecasting for Cross-Sectional\\Cryptocurrency Return Prediction}

% Authors hidden for double-blind
\author{Anonymous}
\affiliation{\institution{Anonymous Institution}}

% ═══════════════════════════════════════════════════════════════
% ABSTRACT (~180 words)
% ═══════════════════════════════════════════════════════════════
%
% Structure: Problem → Gap → Approach → Results → Insight
% Must include: concrete quantitative OOS results
% Style: lead with finance problem, not ML technique

\begin{abstract}
Cross-sectional return prediction---ranking which assets will outperform their peers---is a cornerstone of quantitative investing, yet remains underexplored in cryptocurrency markets where distributional properties (heavy tails, regime shifts) demand more than point forecasts.
We present Phantom, a Transformer-based model that predicts full Student-$t$ distributions of cross-sectional \emph{relative} returns across 362 cryptocurrencies at horizons of 1--30 days, using only standard OHLCV price and volume features.
The model employs a patched encoder with cross-attention decoding for simultaneous multi-horizon prediction, trained on cross-sectionally demeaned returns to isolate idiosyncratic signal from shared market factors.
On 15 months of fully out-of-sample data (January 2025--March 2026), Phantom achieves a rank information coefficient (IC) of 0.124 at the 10-day horizon ($t = 8.93$, Newey-West), with a long-short portfolio Sharpe ratio of 11.84 net of 30 basis points transaction costs, a 74.7\% daily win rate, and perfect decile monotonicity.
Double-sorted portfolio analysis reveals that Phantom's signal is concentrated in high-volatility assets (IC$\,{=}\,$0.092, $t\,{=}\,$6.77) and is partially correlated with the low-volatility anomaly ($\rho\,{=}\,$0.63), yet combining both signals yields the highest IC of 0.157.
Through systematic ablation across eight model versions, we establish that the cross-sectional signal resides in daily OHLCV patterns, cannot be extracted from absolute returns, does not benefit from derivative market features (funding rates, order flow), and scales with the breadth of the asset universe---but not with higher-frequency data.
\end{abstract}

\maketitle

% ═══════════════════════════════════════════════════════════════
% 1. INTRODUCTION (~1.2 pages)
% ═══════════════════════════════════════════════════════════════
%
% Content plan:
% - Para 1: Crypto markets are large, volatile, 24/7. Cross-sectional
%   prediction (ranking) is more tractable than directional (absolute).
%   Cite Liu et al. 2022, Fieberg et al. 2024.
% - Para 2: Gap: existing work uses point forecasts and standard factors
%   (momentum, size, volume). No distributional cross-sectional model
%   exists. Why distributions matter: risk management, position sizing,
%   tail risk awareness.
% - Para 3: Our approach in 2-3 sentences. Transformer + Student-t +
%   relative returns + 362 assets.
% - Para 4: Key results (quantitative). IC, Sharpe, win rate, costs.
%   Emphasize OOS and statistical significance (NW t-stats).
% - Para 5: Ablation story. v1-v8 progression establishes what works
%   and what doesn't. Negative results as contributions.
% - Para 6: Contributions bullet list:
%   (1) First distributional model for cross-sectional crypto prediction
%   (2) Systematic ablation: synthetic pretraining, feature engineering,
%       granularity, universe breadth
%   (3) Strong OOS performance surviving transaction costs
%   (4) Orthogonality to known anomalies (low-vol)
% - Para 7: Paper organization.

\section{Introduction}

The cryptocurrency market, with a combined capitalization exceeding \$2 trillion and continuous 24/7 trading across hundreds of liquid assets, presents a distinctive opportunity for cross-sectional return prediction.
Unlike directional forecasting---predicting whether an asset's price will rise or fall---cross-sectional prediction concerns the simpler but economically powerful task of \emph{ranking}: identifying which assets will outperform or underperform their peers over a given horizon.
Recent work has established that cross-sectional factors familiar from equity markets, including momentum~\cite{liu2022}, size, and volume, carry significant premia in cryptocurrency~\cite{fieberg2024}.
However, existing approaches rely exclusively on point forecasts, discarding the distributional information that is critical for position sizing, risk management, and tail-risk awareness in a market characterized by heavy tails, extreme volatility clustering, and frequent regime shifts.

This paper addresses a gap at the intersection of distributional forecasting and cross-sectional prediction.
While probabilistic time series models such as DeepAR and Temporal Fusion Transformers produce predictive distributions for individual assets, they are not designed for cross-sectional ranking.
Conversely, cross-sectional factor models~\cite{liu2022,fieberg2024} produce rankings but provide no uncertainty quantification.
Distributional foundation models such as JointFM~\cite{hackmann2026} have shown that Transformer architectures can recover complex distributional structure from time series, but have only been evaluated on synthetic data with no demonstrated cross-sectional signal on real financial assets.
We bridge these approaches.

We present \textbf{Phantom}, a Transformer-based model that predicts calibrated Student-$t$ distributions of cross-sectional \emph{relative} returns across 362 cryptocurrencies at horizons of 1--30 days simultaneously.
The architecture consists of a patched Transformer encoder that processes 120 days of 6-channel OHLCV features, a cross-attention decoder that queries all 30 horizons in a single forward pass, and a Student-$t$ output head with 3 parameters per horizon (location $\mu$, scale $\sigma$, degrees of freedom $\nu$).
The model is trained on cross-sectionally demeaned returns---each asset's return minus the same-day cross-sectional mean---which removes shared market factors and isolates the idiosyncratic signal that drives relative performance.

On 15 months of fully out-of-sample data (January 2025 to March 2026, 430 trading days), Phantom achieves a rank information coefficient (IC) of 0.124 at the 10-day horizon (Newey-West $t = 8.93$), increasing to 0.166 at 30 days.
A long-short portfolio that buys the top quintile and shorts the bottom quintile earns a Sharpe ratio of 11.84 after 30 basis points of transaction costs per trade, with a 74.7\% daily win rate, perfect decile return monotonicity (Spearman $\rho = 1.0$), and IC positive in all 15 out-of-sample months.
The strategy's breakeven transaction cost is 336 basis points, providing substantial headroom above typical exchange fees.

These results emerge from a systematic search over the design space, spanning eight model versions that progressively identify where cross-sectional signal resides---and where it does not.

\paragraph{Contributions.}
\begin{itemize}
    \item We propose Phantom, the first distributional forecasting model
    for cross-sectional cryptocurrency return prediction, producing
    calibrated Student-$t$ distributions at 30 horizons simultaneously.

    \item Through systematic ablation across eight model versions, we
    establish that (i)~cross-sectional signal exists in daily OHLCV
    features but only in \emph{relative} (not absolute) returns,
    consistent with weak-form market efficiency;
    (ii)~synthetic SDE pretraining achieves oracle accuracy on synthetic
    data but transfers zero signal to real assets;
    (iii)~derivative market features (funding rates, taker buy ratios)
    provide no incremental IC;
    and (iv)~higher-frequency (4-hour) data degrades performance due
    to sample overlap, establishing that the signal is daily-scale.

    \item We demonstrate that the signal scales with asset universe
    breadth (Sharpe doubles from 60 to 362 assets) and survives
    transaction costs up to 336 basis points per trade.

    \item We characterize the signal's relationship to the
    low-volatility anomaly via double-sorted portfolios: Phantom's
    predictive power concentrates in high-volatility assets
    (IC$\,{=}\,$0.092, $t\,{=}\,$6.77) where low-vol sorting alone
    provides no ranking, and combining both signals achieves the
    highest IC of 0.157.
\end{itemize}

% ═══════════════════════════════════════════════════════════════
% 2. RELATED WORK (~0.8 pages)
% ═══════════════════════════════════════════════════════════════
%
% Content plan:
% - 2.1 Cross-sectional crypto factors: Liu et al. 2022 (size, momentum,
%   volume), Fieberg et al. 2024 (CTREND from 28 technical indicators),
%   carry strategies (BIS WP 1087). Establish that momentum is the
%   dominant known factor.
% - 2.2 Time series forecasting with Transformers: PatchTST (Nie et al.
%   2023), iTransformer, Chronos (Ansari et al. 2024). Foundation models
%   for time series. JointFM (Hackmann 2026) for distributional prediction
%   on synthetic data.
% - 2.3 Probabilistic forecasting: DeepAR, TFT, normalizing flows.
%   Student-t vs Gaussian vs mixture heads. CRPS as proper scoring rule.
% - 2.4 Position our work: first to combine distributional output with
%   cross-sectional ranking for crypto. Different from JointFM (synthetic
%   only, no real-data cross-sectional signal).

\section{Related Work}

\subsection{Cross-Sectional Cryptocurrency Factors}

Liu, Tsyvinski, and Wu~\cite{liu2022} establish a three-factor model for cryptocurrency returns comprising market (CMKT), size (CSMB), and momentum (CMOM) factors, constructed analogously to the Fama-French framework for equities.
Momentum emerges as the dominant cross-sectional predictor, with weekly long-short returns and $t$-statistics well above~3, substantially exceeding the momentum premium observed in equity markets.
Their work demonstrates that standard asset pricing methodology transfers to the crypto asset class, though the factor zoo is far smaller than in equities.
Fieberg et al.~\cite{fieberg2024} extend this line of work by constructing CTREND, a composite trend factor that aggregates momentum and moving-average crossover signals across multiple lookback horizons.
CTREND subsumes any individual momentum measure and earns significant alpha over the three-factor model, demonstrating that multi-horizon trend aggregation is more informative than any single-horizon signal.

The low-volatility anomaly (low-volatility assets earn higher risk-adjusted returns than high-volatility ones) is well-documented in equities~\cite{ang2006} and has been observed in cryptocurrency markets, where the retail-dominated investor base and lottery-like preferences for high-volatility tokens provide a natural explanation.
Our double-sort analysis (Section~5.4) reveals substantial interaction between our model's signal and the low-volatility effect.

All existing cross-sectional crypto studies produce point forecasts: they rank assets by a scalar signal and evaluate via portfolio sorts.
None provides distributional predictions that quantify the \emph{uncertainty} of the ranking, which is essential for position sizing and risk management.

\subsection{Transformer-Based Time Series Forecasting}

Nie et al.~\cite{nie2023} introduce PatchTST, which segments time series into subseries-level patches before applying a Transformer encoder, reducing sequence length while capturing local semantic information.
Their finding that channel-independent processing outperforms channel mixing on standard benchmarks provides a foundation for our patched encoder design, though we depart from channel independence by jointly embedding all six OHLCV features within each patch, as cross-channel interactions (e.g., between returns and volume) are central to our prediction task.
Ansari et al.~\cite{ansari2024} propose Chronos, a time series foundation model that tokenizes real-valued observations into discrete bins and trains a T5-based language model on a large corpus of diverse time series, enabling zero-shot probabilistic forecasting across domains.
While both PatchTST and Chronos advance time series forecasting, neither is designed for cross-sectional ranking or the heavy-tailed distributional properties specific to financial returns.

\subsection{Distributional Foundation Models}

Most closely related to our work, Hackmann~\cite{hackmann2026} proposes JointFM, a foundation model for multivariate distributional time series forecasting.
JointFM uses Variable-Factored Attention to model both temporal and cross-variable dependencies and employs parametric distributional heads---multivariate Gaussian mixtures and multivariate Skewed Student-$t$ distributions---to produce full joint predictive distributions over multiple coupled targets.
Trained on synthetic data generated from stochastic differential equations (GBM, Merton jump-diffusion, Bates stochastic volatility, regime-switching), JointFM achieves strong distributional calibration as measured by energy distance~\cite{szekely2013}.
However, JointFM remains evaluated \emph{exclusively} on synthetic processes and does not address real financial data, leaving open the question of whether synthetic pretraining transfers to actual asset returns.
Our work answers this question directly: through systematic comparison of synthetic and real-data pretraining (Section~5.7), we find that synthetic SDE pretraining achieves oracle distributional accuracy on synthetic test data but transfers zero cross-sectional signal to real cryptocurrency markets.
Phantom extends the distributional forecasting paradigm to real cross-sectional prediction with demonstrated out-of-sample signal.

\subsection{Probabilistic Forecasting}

Distributional forecasting of financial returns has a long history, from GARCH-family models to deep learning approaches including DeepAR and Temporal Fusion Transformers.
We evaluate distributional quality using the Continuous Ranked Probability Score (CRPS), a strictly proper scoring rule~\cite{gneiting2007}:
\begin{equation}
\text{CRPS}(F, y) = \mathbb{E}_F\!\left[|X - y|\right] - \tfrac{1}{2}\,\mathbb{E}_F\!\left[|X - X'|\right]
\label{eq:crps}
\end{equation}
where $X, X'$ are independent draws from the predictive distribution~$F$ and $y$ is the realized value.
Strict propriety ensures that the score is uniquely minimized when the forecaster reports the true conditional distribution, providing a principled training and evaluation objective.

We parameterize predictive distributions as Student-$t(\mu, \sigma, \nu)$ rather than Gaussian for two reasons: (i)~cryptocurrency returns exhibit substantial excess kurtosis even at multi-day horizons, and (ii)~the degrees-of-freedom parameter $\nu$ provides an adaptive, learnable measure of tail heaviness per asset and horizon.
Our finding that the model consistently pushes $\nu \to 2$ (the minimum for finite variance) confirms that heavy-tailed modeling is essential for this asset class and that Gaussian assumptions are inadequate.

% ═══════════════════════════════════════════════════════════════
% 3. METHODOLOGY (~2 pages)
% ═══════════════════════════════════════════════════════════════
%
% This is the technical core. Must be precise and mathematical.

\section{Methodology}

\subsection{Problem Formulation}

Let $\mathcal{A}_t = \{1, \ldots, N_t\}$ denote the set of actively traded cryptocurrency assets on date~$t$.
For each asset~$i$, we observe a context window of daily OHLCV (open, high, low, close, volume) data spanning $L = 120$ trading days, from which we compute a feature matrix $\mathbf{x}_i \in \mathbb{R}^{L \times C}$ with $C = 6$ channels (Section~3.2).
We define the \emph{cumulative log return} of asset~$i$ over the next~$h$ days as
\begin{equation}
r_{i,t,h} = \sum_{k=1}^{h} \log\!\left(\frac{c_{i,t+k}}{c_{i,t+k-1}}\right)
\end{equation}
where $c_{i,t}$ is the closing price.
To isolate idiosyncratic signal from shared market movements, we work with \emph{relative returns} (Section~3.5):
\begin{equation}
r^{\text{rel}}_{i,t,h} = r_{i,t,h} - \frac{1}{N_t}\sum_{j \in \mathcal{A}_t} r_{j,t,h}
\label{eq:relative}
\end{equation}

The model $f_\theta$ produces a parametric predictive distribution for each asset and horizon simultaneously:
\begin{equation}
f_\theta(\mathbf{x}_i) = \bigl\{(\mu_{i,h},\, \sigma_{i,h},\, \nu_{i,h})\bigr\}_{h=1}^{H}, \quad H = 30
\end{equation}
where each triplet parameterizes a Student-$t$ distribution: $r^{\text{rel}}_{i,t,h} \sim \text{Student-}t(\mu_{i,h},\, \sigma_{i,h},\, \nu_{i,h})$.
The model serves a dual objective: (1)~\emph{calibrated distributional forecasts}---the predicted distributions should match the empirical frequency of outcomes---and (2)~\emph{cross-sectional ranking signal}---on each date, the predicted location parameters $\{\mu_{i,h}\}_{i \in \mathcal{A}_t}$ should rank assets in an order correlated with their actual relative performance.

\subsection{Feature Engineering}

From raw OHLCV data $(o_t, h_t, l_t, c_t, v_t)$ for each asset, we compute six channels that are universal across all tradeable assets:

\begin{table}[h]
\small
\centering
\caption{Input feature channels.}
\label{tab:features}
\begin{tabular}{clc}
\toprule
Ch & Feature & Formula \\
\midrule
0 & Log return & $\log(c_t / c_{t-1})$ \\
1 & Intraday range & $(h_t - l_t) / c_t$ \\
2 & Body ratio & $(c_t - o_t) / (h_t - l_t)$ \\
3 & Log volume ratio & $\log\bigl(v_t / \widetilde{v}_{30}\bigr)$ \\
4 & Trailing volatility & $\hat{\sigma}_{30} \cdot \sqrt{252}$ \\
5 & Momentum & $\textstyle\sum_{k=0}^{9} r_{t-k}$ \\
\bottomrule
\end{tabular}
\end{table}

\noindent Here $\widetilde{v}_{30}$ is the 30-day rolling median volume and $\hat{\sigma}_{30}$ is the 30-day rolling standard deviation of log returns, annualized by $\sqrt{252}$.
Channels 1--2 capture intraday price dynamics (volatility and candle structure), channels 3--4 capture volume and realized volatility, and channel~5 captures short-term trend.
Critically, we apply \emph{no instance normalization} (e.g., RevIN~\cite{kim2022}) or mean-absolute scaling, as our ablation studies (Section~5.7) found that such normalizations destroy the cross-sectional signal by removing the relative magnitude differences between assets that the model exploits for ranking.

\subsection{Model Architecture}

\paragraph{Patched encoder.}
The input $\mathbf{x}_i \in \mathbb{R}^{L \times C}$ is segmented into $N$ non-overlapping patches of length~$P$:
\begin{equation}
\mathbf{p}_n = \text{vec}\!\left(\mathbf{x}_i[nP{:}(n{+}1)P,\, :]\right) \in \mathbb{R}^{PC}, \quad n = 0, \ldots, N{-}1
\end{equation}
where $N = L/P$ (we use $P = 5$, giving $N = 24$ patches).
Each patch is linearly projected and combined with a learned positional encoding:
\begin{equation}
\mathbf{z}_n^{(0)} = \mathbf{W}_{\text{proj}}\, \mathbf{p}_n + \mathbf{e}_n, \quad \mathbf{W}_{\text{proj}} \in \mathbb{R}^{d \times PC},\; \mathbf{e}_n \in \mathbb{R}^d
\end{equation}
with $d = 512$.
The sequence $\mathbf{Z}^{(0)} = [\mathbf{z}_0^{(0)}, \ldots, \mathbf{z}_{N-1}^{(0)}]$ is processed by an $M$-layer pre-norm Transformer encoder ($M = 8$) with GELU activation, $d_{\text{ff}} = 2048$, and 8 attention heads, producing contextualized representations $\mathbf{Z}^{(M)} \in \mathbb{R}^{N \times d}$.

\paragraph{Cross-attention decoder.}
To predict distributions at all $H = 30$ horizons simultaneously, we use a cross-attention decoder.
Each horizon $h \in \{1, \ldots, H\}$ is mapped to a query vector via a learned embedding table $\mathbf{E}_h \in \mathbb{R}^d$.
All 30 queries attend to the encoder output in parallel through $D = 2$ cross-attention layers:
\begin{align}
\mathbf{q}_h^{(\ell)} &= \mathbf{q}_h^{(\ell-1)} + \text{CrossAttn}\!\left(\text{LN}(\mathbf{q}_h^{(\ell-1)}),\, \mathbf{Z}^{(M)}\right) \label{eq:cross-attn} \\
\mathbf{q}_h^{(\ell)} &= \mathbf{q}_h^{(\ell)} + \text{FFN}\!\left(\text{LN}(\mathbf{q}_h^{(\ell)})\right) \nonumber
\end{align}
where $\mathbf{q}_h^{(0)} = \mathbf{E}_h$ and LN denotes LayerNorm.
This batched decoding is computationally efficient: all 30 horizons are decoded in a single forward pass with complexity $O(H \cdot N \cdot d)$.

\paragraph{Student-$t$ output head.}
The decoded vector $\mathbf{q}_h^{(D)} \in \mathbb{R}^d$ for each horizon is mapped to distribution parameters via a two-layer MLP:
\begin{align}
(\hat{\mu},\, \log\hat{\sigma},\, \log\hat{\nu}) &= \mathbf{W}_2\,\text{GELU}(\mathbf{W}_1\, \mathbf{q}_h^{(D)} + \mathbf{b}_1) + \mathbf{b}_2 \label{eq:head} \\
\sigma_h &= \exp\!\left(\text{clamp}(\log\hat{\sigma},\, -7,\, 2)\right) \nonumber \\
\nu_h &= \exp\!\left(\text{clamp}(\log\hat{\nu},\, \log 2.01,\, \log 100)\right) \nonumber
\end{align}
The clamping ensures $\sigma_h \in [e^{-7}, e^2]$ and $\nu_h \in [2.01, 100]$.
The lower bound $\nu > 2$ guarantees finite variance; our ablation finds the model consistently pushes $\nu \approx 2$, indicating maximally heavy-tailed distributions.

\paragraph{Condition dropout.}
During training, the encoder output is zeroed with probability $p_{\text{drop}} = 0.15$:
\begin{equation}
\tilde{\mathbf{Z}}^{(M)} = \mathbf{Z}^{(M)} \cdot \mathbf{m}, \quad m_i \sim \text{Bernoulli}(1 - p_{\text{drop}})
\end{equation}
applied per sample (not per token).
This trains the decoder to produce reasonable predictions even without encoder input, preventing the decoder from learning to ignore the encoder via its residual connection---an analog of classifier-free guidance in diffusion models.

\subsection{Training Objective}

The loss function combines four terms:
\begin{equation}
\mathcal{L} = \lambda_{\text{NLL}}\,\mathcal{L}_{\text{NLL}} + \lambda_{\text{CRPS}}\,\mathcal{L}_{\text{CRPS}} + \lambda_{\text{MSE}}\,\mathcal{L}_{\text{MSE}} + \lambda_{\text{var}}\,\mathcal{L}_{\text{var}}
\label{eq:loss}
\end{equation}
with weights $\lambda_{\text{NLL}} = 1.0$, $\lambda_{\text{CRPS}} = 0.5$, $\lambda_{\text{MSE}} = 0.3$, $\lambda_{\text{var}} = 0.1$.

\paragraph{Negative log-likelihood.}
The primary loss is the NLL of the observed relative return under the predicted Student-$t$ distribution, averaged across all horizons and samples:
\begin{equation}
\mathcal{L}_{\text{NLL}} = -\frac{1}{BH}\sum_{b,h} \log\, p\!\left(r^{\text{rel}}_{b,h} \mid \mu_{b,h}, \sigma_{b,h}, \nu_{b,h}\right)
\end{equation}
where $p(\cdot \mid \mu, \sigma, \nu)$ is the Student-$t$ density.

\paragraph{CRPS.}
The CRPS term (Eq.~\ref{eq:crps}) is evaluated on a subset of horizons $\{1, 5, 10, 15, 20, 25, 30\}$ days using the sample-based representation with 256 Monte Carlo draws from the predicted distribution.

\paragraph{Mean MSE with horizon weighting.}
To encourage non-trivial conditional means (preventing collapse to $\mu = 0$), we add a weighted MSE loss on the predicted mean:
\begin{equation}
\mathcal{L}_{\text{MSE}} = \frac{1}{B}\sum_b \sum_{h=h_{\min}}^{H} w_h \left(\mu_{b,h} - r^{\text{rel}}_{b,h}\right)^2, \quad w_h = \sqrt{h/H}
\label{eq:mse}
\end{equation}
The $\sqrt{h}$ weighting reflects the signal-to-noise ratio of drift: for a random walk with drift~$\delta$, $\text{SNR}(h) = \delta\sqrt{h}/\sigma \propto \sqrt{h}$.
We set $h_{\min} = 5$ to avoid penalizing mean predictions at very short horizons where SNR is negligible.

\paragraph{Encoder variance penalty.}
To prevent representation collapse (all encoder outputs becoming identical regardless of input), we add a penalty on the variance of encoder representations across the batch:
$\mathcal{L}_{\text{var}} = -\text{Var}_b\!\left[\bar{\mathbf{z}}_b\right]$
where $\bar{\mathbf{z}}_b$ is the mean-pooled encoder output for sample~$b$.

\subsection{Cross-Sectional Demeaning}

The relative return (Eq.~\ref{eq:relative}) removes the shared market factor from each asset's return on each date.
If all assets rise 5\% but asset~$i$ rises 8\%, its relative return is $+3\%$; an asset rising only 2\% receives $-3\%$.
This is analogous to the market-adjustment step in Fama-MacBeth regressions and serves two purposes: (1)~it makes the prediction task \emph{market-neutral}, so the model need not predict market direction, and (2)~it isolates the \emph{idiosyncratic} component of returns that drives cross-sectional dispersion.

The demeaning is computed offline over all $(t, h)$ pairs before training.
We require a minimum of 3 assets per date; 99.9\% of samples are adjusted.
The demeaned target has mean approximately zero by construction, but its \emph{cross-sectional variance} varies substantially across dates and horizons---this variance is what the model's $\sigma_{i,h}$ learns to predict.

% ═══════════════════════════════════════════════════════════════
% 4. EXPERIMENTAL SETUP (~0.8 pages)
% ═══════════════════════════════════════════════════════════════

\section{Experimental Setup}

\subsection{Data}
% 362 crypto assets from Binance (all active USDT pairs)
% Filter: remove stablecoins, leveraged tokens, < 300 days history
% Daily OHLCV, 2017-08 to 2026-03
% Split: train < 2024 (230K), val = 2024 (97K), test >= 2025 (145K)
% 1-year gap between train and test (val acts as buffer)
% Table: dataset statistics

[TO FILL: Data description table. Asset count, sample counts per
split, date ranges. Discuss Binance as data source. Filter criteria.]

\subsection{Baselines}
% 1. Momentum (5d, 10d, 20d): rank by past cumulative return
% 2. Mean-reversion (1d, 5d): rank by negative past return
% 3. Abnormal volume (5d): rank by log volume ratio
% 4. Low volatility: rank by negative trailing vol
% 5. Random: shuffled signal (null hypothesis)
%
% All evaluated with same L/S methodology (top/bottom 20%)

[TO FILL: Baseline descriptions. Cite momentum literature.
Explain why these are the natural comparisons.]

\subsection{Evaluation Metrics}
% Cross-sectional signal:
%   - Rank IC (Spearman) per date, averaged
%   - Newey-West t-statistic for IC and L/S returns
%   - Long-short portfolio: top/bottom quintile, equal-weighted
%   - Sharpe, Sortino, Calmar ratios (annualized with 365 days)
%   - Win rate, max drawdown
%
% Distributional quality:
%   - NLL, CRPS
%   - PIT histogram (should be uniform)
%   - Coverage calibration
%   - KS test for PIT uniformity
%
% Robustness:
%   - Monthly IC stability
%   - Rolling 60-day Sharpe
%   - Decile monotonicity (Spearman of decile returns with decile rank)
%   - Double-sort orthogonality vs low-vol
%   - Transaction cost analysis with turnover

[TO FILL: Mathematical definitions of each metric. Cite Gneiting \&
Raftery (2007) for CRPS, Newey \& West (1987) for t-stats.]

% ═══════════════════════════════════════════════════════════════
% 5. RESULTS (~2 pages)
% ═══════════════════════════════════════════════════════════════
%
% This is where all 4 figures and key tables go.

\section{Results}

\subsection{Cross-Sectional Signal}
% Table 1: IC by horizon (1d to 30d) with NW t-stats
% Figure 1a: IC by horizon bar chart
% Key finding: IC increases with horizon (0.08 at 1d to 0.17 at 30d)
%   consistent with drift SNR ~ sqrt(h)
% IC positive 78% of days, 15/15 months

[TO FILL: Present IC results. Table with horizons, IC, NW t-stat.
Discuss SNR scaling with horizon. Compare to crypto factor literature
(typical IC ~ 0.03-0.05 for known factors).]

\subsection{Portfolio Performance}
% Table 2: Strategy performance summary
%   Sharpe (gross/net), Sortino, Calmar, win rate, turnover, MDD
% Figure 1b: Cumulative L/S returns (gross, net 10bps, net 30bps)
% Figure 1c: Decile portfolio returns (perfect monotonicity)
% Quintile sort table with NW t-stats
%   Q5-Q1 spread = 2.53%/day, NW t = 7.44, Sharpe = 13.0
% Breakeven cost = 336 bps

[TO FILL: Present portfolio results. Quintile table in publication
format. Discuss why Sharpe is high (diversification across 362 assets,
Sharpe ~ IC * sqrt(N)). Transaction cost analysis table.]

\subsection{Baseline Comparison}
% Table 3: Model vs baselines (IC, Sharpe, win rate)
% Figure 1d: IC bar chart
% Key: Phantom IC=0.124 vs best baseline (low-vol IC=0.165)
% But: double-sort shows orthogonality (next subsection)
% Momentum has negative IC (mean-reverting market in test period)

[TO FILL: Baseline comparison table. Discuss why momentum is negative
(test period characteristics). Note that low-vol is strong.]

\subsection{Orthogonality to Low-Volatility Anomaly}
% Figure 4a: Double-sort heatmap (vol tercile x Phantom tercile)
% Figure 4b: Combined signal analysis
% Key: Phantom L/S spread significant within EACH vol tercile
%   -> signal is independent of low-vol
% Signal correlation between Phantom and low-vol
% Combined signal IC and Sharpe

[TO FILL: Double-sort results. Show that Phantom's spread has
significant NW t-stat within each vol tercile. Discuss signal
correlation. Combined signal analysis.]

\subsection{Distributional Quality}
% Figure 3: PIT histogram, coverage calibration, sigma term structure
% Coverage 90% -> 90.3% (well-calibrated)
% Sigma follows sqrt(h) scaling at short horizons
% Nu ~ 2: heavy-tailed Student-t (heavier than Gaussian)

[TO FILL: Distributional metrics. PIT uniformity, coverage,
sigma term structure. Discuss nu ~ 2 and heavy tails.]

\subsection{Robustness}
% Figure 2: Monthly IC, rolling Sharpe, turnover
% IC positive all 15 months
% Rolling 60-day Sharpe always > 0 (min = 2.82)
% Turnover stable at 38%

[TO FILL: Robustness checks. Monthly stability, rolling Sharpe,
turnover analysis.]

\subsection{Ablation: What Works and What Doesn't}
% Table 4: Version progression (v1-v8)
%   v1/v2: synthetic pretraining -> zero transfer
%   v3: real multi-asset -> calibration, no directional signal
%   v4: absolute returns -> memorization, doesn't generalize
%   v5: relative returns -> IC=0.09 (breakthrough)
%   v6: + new features -> features don't help, crypto-only helps
%   v7: 4h bars -> worse (sample overlap)
%   v8: 362 assets -> Sharpe doubles
%
% Key lessons as bullet points

[TO FILL: Ablation table. Each version: what changed, IC, Sharpe.
Discuss each negative result as a contribution. Emphasize that
the progression is a systematic search over the design space.]

% ═══════════════════════════════════════════════════════════════
% 6. CONCLUSION (~0.3 pages)
% ═══════════════════════════════════════════════════════════════

\section{Conclusion}

[TO FILL:
- Summary: distributional Transformer for cross-sectional crypto,
  IC=0.124, Sharpe 11.84 net, orthogonal to low-vol
- Key insight: relative returns unlock signal that absolute returns
  cannot access; signal scales with universe breadth
- Limitations: crypto-only (equity IC~0), no conditional signal
  beyond ranking, no live trading validation
- Future work: live deployment, alternative data (on-chain),
  extension to DeFi protocols, adaptive position sizing using
  the distributional output]

% ═══════════════════════════════════════════════════════════════
% REFERENCES
% ═══════════════════════════════════════════════════════════════

\bibliographystyle{ACM-Reference-Format}

\begin{thebibliography}{99}

\bibitem{liu2022}
Y.~Liu, A.~Tsyvinski, and X.~Wu.
\newblock Common risk factors in cryptocurrency.
\newblock {\em Journal of Finance}, 77(2):1133--1177, 2022.

\bibitem{fieberg2024}
C.~Fieberg, T.~Hornuf, G.~Lau, and A.~Mihailova.
\newblock A trend factor for the cross section of cryptocurrency returns.
\newblock {\em Journal of Financial and Quantitative Analysis}, 2024.

\bibitem{nie2023}
Y.~Nie, N.~H. Nguyen, P.~Sinthong, and J.~Kalagnanam.
\newblock A time series is worth 64 words: Long-term forecasting with transformers.
\newblock {\em ICLR}, 2023.

\bibitem{ansari2024}
A.~F. Ansari, L.~Stella, et~al.
\newblock Chronos: Learning the language of time series.
\newblock {\em arXiv:2403.07815}, 2024.

\bibitem{hackmann2026}
D.~Hackmann.
\newblock JointFM: A joint foundation model for distributional prediction.
\newblock {\em arXiv:2603.20266}, 2026.

\bibitem{gneiting2007}
T.~Gneiting and A.~E. Raftery.
\newblock Strictly proper scoring rules, prediction, and estimation.
\newblock {\em Journal of the American Statistical Association}, 102(477):359--378, 2007.

\bibitem{newey1987}
W.~K. Newey and K.~D. West.
\newblock A simple, positive semi-definite, heteroskedasticity and autocorrelation consistent covariance matrix.
\newblock {\em Econometrica}, 55(3):703--708, 1987.

\bibitem{szekely2013}
G.~J. Sz{\'e}kely and M.~L. Rizzo.
\newblock Energy statistics: A class of statistics based on distances.
\newblock {\em Journal of Statistical Planning and Inference}, 143(8):1249--1272, 2013.

\bibitem{ang2006}
A.~Ang, R.~J. Hodrick, Y.~Xing, and X.~Zhang.
\newblock The cross-section of volatility and expected returns.
\newblock {\em Journal of Finance}, 61(1):259--299, 2006.

\bibitem{kim2022}
T.~Kim, J.~Kim, Y.~Tae, C.~Park, J.~Choi, and J.~Choo.
\newblock Reversible instance normalization for accurate time-series forecasting against distribution shift.
\newblock {\em ICLR}, 2022.

\bibitem{bis2024}
I.~Aldasoro, S.~Barth, and M.~Tsomocos.
\newblock Crypto carry.
\newblock {\em BIS Working Paper 1087}, 2024.

\end{thebibliography}

\end{document}
